---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# genero

<!-- badges: start -->
<!-- badges: end -->

The goal of `genero` is to ease data cleaning when it comes to augment data with "gender" or sex assigned at birth. It uses lists of common names in Spanish and Portuguese with frequent gender assignments. Therefore it has many limitations and should be used with care. This package is provided to study large populations and not individual records.

For better methods for estimating gender in English and other languages please use the [gender package](https://github.com/ropensci/gender/)

This package does not take into account non-binary gender identification as of this moment. We apologize for that, we expect to have more sensical approaches in the near future. Any ideas or feedback on how to make it more inclusive are more than welcome.

Use this package with care and make sure it is not used for any type of discrimination. 

Make sure to check the results provided by `genero` by other means, like manual inspection.

## Installation

You can install the released version of genero from [CRAN](https://CRAN.R-project.org) (SOOOOOON) with:

``` r
install.packages("genero")
```

And the development version with 

``` r
devtools::install_github("datasketch/genero")
```

## Example

This is a basic example which shows you how to estimate gender from names:

```{r example}
library(genero)

names <- c("Juan", "Pablo", "Camila", "Mariana")
genero(names)

```

For Portuguese use the `lang = "pt"` parameter. Don't like the results English? Me neither, simply provide a named vector to rename results.


```{r example-2}
names <- c("Gabriela", "Ina","Luiz", "Edson")
genero(names, lang = "pt", result_as = c(male = "Homem", female = "Mulher"))
```

It also works with data frames and composed names, and even last names. The `genero` tries to guess whats the column with the names, if it finds two columns with names, the first one will be used.

```{r example-3}
names <- c("Juan Álvarez", "Juan Maria García", "Maria José", "José María", "María Rodríguez")
age <- c(23, 43, 56, 67, 24)

d <- data.frame(names = names, age = age, stringsAsFactors = FALSE)
genero(names)
```




